Dual host-parasite transcriptomes of apicomplexan Eimeria falciformis and its natural mouse host

### 2016-08-15 manuscript/bmc_article.tex
At the stage of merging patterns over samples and GO-Terms : Handing over to Totta!

Statistical testing for differetial expression between infected and
uninfected mice, revealed changes in mRNA abundance becoming more
pronounced (involving more genes) at later post infection (especially
at 7dpi; Table 1).

@@

The changes in early infection are ...

Those changes are similarly pronounce in the immunocompetent strains
and RAG-/- immunodefficient animals.

The ponounced changes later in infection are is likely reflecting the
onset of an adaptive immune response.

@@

Table 2 lists pairwise sample comparisons and number of genes with
differently abundant mRNAs per comparison."NMRI" followed by number
indicates day post infection (e.g. NMRI 3 = NMRI mouse on day 3
p.i.). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as
implemented in edgeR are included. NAs are missing samples or not
applicable for the species. NA* is due to missing NMRI day 0 sample
from first infection. For \textit{E. falciformis} no mRNAs are
significantly differently abundant between first and second infection,
whereas in mouse there are some differences, especially on day 7
p.i. However, upon hierarchical clustering of these genes (union of
500 with lowest FDR), samples do not cluster according to first/second
infection (Figure SX ?). In comparisons between mouse strains, 22
genes are detected to be different in the parasite data. The
difference is however not between Rag1-/- mice as expected, but
between NMRI and C57BL/6. For mouse....

We applied hierarchical clustering on \textit{E. falciformis} mRNAs
which were differently abundant at different timepoints p.i.,
including comparisons with oocysts and sporozoites (union of 500 with
lowest FDR). Samples from day 7 p.i. form a separate cluster (NMRI
mice only), not distinguishing between first and challenge
infection. For these samples, the pattern is most pronounced in gene
cluster 1 (up), and 4 and 6 (down). Within cluster 1, the second
replicate from the challenge infection has a deviant profile (compare
mouse data in Figure 2). Distinct groups of genes also define
sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6
down). mRNA profiles on days 3 and 5 p.i. from all three mouse strains
cluster together. These samples are distinct from oocysts, NMRI day 7
p.i., and sporozoites. Sporozoites cluster closer with days 3 and 5
than with oocysts or day 7.

>>>>

Statistical testing for differetial expression between infected and
uninfected mice, revealed changes in mRNA abundance becoming more
pronounced (involving more genes) at later post infection (especially
at 7dpi; Table 2).

@@

The ponounced changes later in infection are reflecting the expected
onset of an adaptive immune response. This is underlined by GO-terms
such as ``antigen binding'' and ``immunoglobulin receptor binding''
(molecular function) ``immune system process'' and ``adaptive immune
response'' (biological process) highly enriched in Cluster 5 of figure
2c.

Particular ``regulation of cell death'' and ``regulation of apoptotic
process'' (BP)

``UDP-N-acetylglucosamine-lysosomal-enzyme'' (MF)

%% This might be relevant: The DMAP interaction domain of
%% UDP-GlcNAc:lysosomal enzyme
%% N-acetylglucosamine-1-phosphotransferase is a substrate recognition
%% module. Catalysis of the reaction: UDP-N-acetyl-D-glucosamine +
%% lysosomal-enzyme D-mannose = UMP + lysosomal-enzyme
%% N-acetyl-D-glucosaminyl-phospho-D-mannose

General changes upon infection are a downregulation of genes involved
``lipid modification'' or more detailed ``phosphatidylinositol
phosphorylation''. These changes are similarly pronounce in the
immunocompetent strains and RAG-/- immunodefficient animals (cluster 3
of fibure 2c) (see supplementary table X).  The presence of these
changes immunocompromized animals invite the speculation that those
genes are independent of a potential later adaptive immune response.

Changes not detectable in RAG-/- immunodefficient animals (cluster 3
of fibure 2c) but otherwise observed in general infection are
characterized by and enrichment to genes involved in ``positive
regulation of calcineurin-NFAT'', ``blood coagulation'' and ``inositol
phosphate-mediated signaling'' (see supplementary table X). The
abscence of these changes in RAG-/- mice allow us to speculate that
those genes might be involved in a very early onlset of adaptive
immune response.

%% Go on here if you can...

We dedtected further differences between infections of naive and
challenge infected mous hosts , especially at 7 dpi. These genes are
characterized by...
%% We should't limit ourselfs to testing the clusters only! But
%% discuss the cluster first if you got difficulties coding the
%% enrichment tests.

@@

Table 2 lists pairwise sample comparisons and number of genes with
differently abundant mRNAs per comparison.  For
\textit{E. falciformis} no mRNAs are significantly differently
abundant between first and second infection.

Changes throughout the intr-host lifecycle can be clearly categorized
in an early (3 and 5 dpi) and late (7dpi) phase. Sporozoites show
expression profiles similar more similar to early infection stages
while oocysts show distinctly different mRNA abundance profile (MDS
plot .

We applied hierarchical clustering on \textit{E. falciformis} mRNAs
which were differently abundant at different dpi, including
comparisons with oocysts and sporozoites.

Samples from 7dpi form a separate cluster (NMRI mice only), not
distinguishing between first and challenge infection.

%% Mix it all well into the GO-term discussion! GO-term inferred
%% biiology needs to be reported in connection to clutering-inferred
%% ``sample biology''.

%% For these samples, the pattern is most pronounced in gene cluster 1
%% (up), and 4 and 6 (down). Within cluster 1, the second replicate from
%% the challenge infection has a deviant profile (compare mouse data in
%% Figure 2). Distinct groups of genes also define sporozoites (cluster 5
%% up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days
%% 3 and 5 p.i. from all three mouse strains cluster together. These
%% samples are distinct from oocysts, NMRI day 7 p.i., and
%% sporozoites. Sporozoites cluster closer with days 3 and 5 than with
%% oocysts or day 7.
